REPRODUCIBLE RESEARCH: ASSIGNMENT 1  



```{r setoptions, echo=FALSE, results='hide',message=FALSE}
library(ggplot2)
library(Hmisc)
options(warn=0)
```

DOWNLOAD AND LOAD ACTIVITY DATA 
```{r, echo=TRUE}
if(!file.exists('activity.csv')){
  unzip('activity.zip')
}
data <- read.csv('activity.csv')

str(data)
head(data)
```

FORMAT DATE 
```{r, echo=TRUE}
data$date <- as.Date(data$date)
head(data)
class(data$date)
```


CALCULATE TOTAL STEPS TAKEN PER DAY:

```{r, echo=TRUE,warning=FALSE}
steps_day <- tapply(data$steps, data$date, sum)
qplot(steps_day, xlab="Total steps per day", ylab="Frequency", binwidth=500)
```
   

CALCULATE THE MEAN AND MEDIAN OF THE DATA SET:
```{r, echo=TRUE}
Daily_steps <- tapply(data$steps, data$date, sum, na.rm=TRUE)      
mean_step <- mean(Daily_steps)
mean_step
median_step <- median(Daily_steps)
median_step
```



CALCULATE THE MAXIMUM, 5-MINUTE INTERVAL:

```{r, echo=TRUE,warning=FALSE}
steps_interval <- aggregate(steps ~ interval, data, mean,rm.na=TRUE)
names(steps_interval)[2] <- "MeanSteps"

plot(steps_interval$interval,
     steps_interval$MeanSteps,
     type = "l",
     main= "TIME-SERIES: Average of 5 minute Intervals, Averaged Accross all Days",
     xlab="5-Minute Intervals",
     ylab="Average Steps Taken")
```


WHAT INTERVAL CONTAINS THE MAX MEAN NUMBER OF STEPS?

```{r, echo=TRUE, warning=FALSE}
Max_Steps <- steps_interval[steps_interval$MeanSteps==max(steps_interval$MeanSteps),]
Max_Steps
```



IMPUTE THE MISSING VALUE

```{r, echo=TRUE}
missing_values <- length(which(is.na(data$steps)))
```

CREATE AN IMBUTED DATA SET, REPLACING NA VALUES WITH MEAN. 

```{r, echo=TRUE}
NA_amount <- length(which(is.na(data$steps)))
data_imputed <- data
data_imputed$steps <- impute(data$steps, fun=mean)
```

CREATE A HISTOGRAM OF IMBUTED DATA:
     
```{r, echo=TRUE}
steps_day_imputed <- tapply(data_imputed$steps, data_imputed$date, sum)
qplot(steps_day_imputed, xlab="Total steps per day (Imputed)", ylab="Frequency", binwidth=500)

#CALCULATE MEAN AND MEDIAN OF IMBUTED DATA:
Daily_steps_imbuted <- tapply(data_imputed$steps, data_imputed$date, sum) 
steps_day_imputed_mean <- mean(Daily_steps_imbuted)
steps_day_imputed_mean
steps_day_imputed_median <- median(Daily_steps_imbuted)
steps_day_imputed_median
```

Comparing activity patterns between weekends and weekdays

```{r, echo=TRUE}
week_data <- data_imputed
week_data$week <- weekdays(week_data$date)
week_data$type <- as.factor(c("weekend", "weekday"))
week_data[week_data$week == "Sunday" | week_data$week == "Saturday" ,5]<- factor("weekend")
week_data[!(week_data$week == "Sunday" | week_data$week == "Saturday"),5 ]<- factor("weekday")

str(week_data)
```

Make a time series panel plot of the average 5-minute interval of steps taken.
Average these accross weekends and weekdays

Createa variable containing the aggregate mean steps taken per 5 minute inteval. 
```{r, echo=TRUE}
interval_week_mean <- aggregate(steps ~ interval + type, week_data, mean)
names(interval_week_mean)[3] <- "mean_steps"
str(interval_week_mean)


par(mar=c(1,1,1,1))
plot(y=interval_week_mean$mean_steps,
     x=interval_week_mean$interval,
     type = "l",
     main = "TIME-SERIES",
     sub = "Mean of 5-minute intervals, averaged accross weekends and weekdays",
     xlab="5-Minute Intervals",
     ylab="Average Steps Taken")
```


